---
title: "The Social Norms behind Inaction and Why Remaining Stagnant Can Make You Regret"
author: Janssen Myer Rambaud
thanks: "Code and data are available at: https://github.com/foreverwoods/The-Normality-of-Action-Inaction-and-How-They-Interact-with-Regret; Replication on Social Science Reproduction platform available at: "
date: today
date-format: long
abstract: "The general idea regarding the concept of regret can be summarized with the famous phrase of 'Don't do something you may regret'. Through experiments conducted by Gilad Feldman, where participants were given a scenario in which they responded with a scaled level of how they perceived regret in said scenario, there was evidence that hinted at a correlation between inaction, perceived regret and societal norms. As humans we tend to be cautious in our actions and at times be too hesitant in making a move in which we will later regret. This paper aims to address this rather undiscussed topic and why understanding more about societal norms and how it affects our inaction can help achieve greater understanding about why we feel regret."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(knitr)
```

# Introduction

As humans, we are all familiar with the general feeling of regret. For the most part these regrets are caused by mistakes on our part, but other times, they can also be the result of not taking action. The saying, 'hindsight is 20/20', is one specific example that can be attributed to this feeling of regret. Before diving deeper into this topic, it is best to begin with a definition of regret and inaction. According to psychology and behaviour scholar Marcel Zeelenberg, regret is defined as, "the negative, cognitively based emotion that we experience when realizing or imagining that our present situation would have been better had we acted differently" (Roos 2023) While inaction can be defined simply as not taking action or not acting upon something. This here drives a question of why, as humans, feel that way. Researcher Gilad Feldman interprets this feeling to be influenced by normality, specifically, societal norms in general (Feldman 2019).

Diving deeper into this topic, it is actually possible to look into what can attribute to a 'quantifiable' amount of regret. This can be called 'perceived regret'. While it is true that every person will feel a different level of regret, and even then, it would be difficult to express just how much 'regret' one has, it would be wise to gather a fairly larger group of people to tackle the possible problem of bias. As such, multiple experiments were conducted by Feldman under a simple question and rating system. This is where the calculation of perceived regret is conducted. The participants were provided scenarios and asked questions pertaining a particular party in the scenario and what they believe their 'perceived regret' is.

By presenting a fixed scenario in these experiments, each participant can answer in a more rational, yet personalized manner. If you were to ask them to recall their own experiences for the sake of gathering a variety of responses/scenarios, the data could possibly be all over the place. Each person has their own experiences, some consisting of extreme regret, while others do not. These are problems that could possibly influence the data in some manner. It is important to consider this idea of inaction because it is something that we do in our lives, and something we will continue to do as we go on in life. I am looking to validate Feldman's claims that normality plays a crucial role in impacting the perception of regret. The research was studied on undergraduate Hong Kong students as well as hired MTurk participants. The estimand in question is the perceived regret that stems from inaction and how societal norms come into play. This paper aims to follow a reproduction of Feldman's research and findings, but dwell deeper onto the side of inaction and norm theory rather than the comparison between action and inaction.

# Data {#sec-data}

Using R from @citeR as the programming language, we can conduct research based on the research provided by Gilad Feldman in his journal article, "What is normal? Dimensions of action-inaction normality and their impact on regret in the action-effect" (@feldmanRegrets) as well as the data from the experiments (@feldmanDataset), we can dive deeper into the concept of inaction. Other articles were also used, such as @regretDefinition in order to give a sound understanding to our topic.

From here, the usage of two datasets from the original usage will be put to work. In a successful attempt to recreate the original table, I was able to clean the raw data to align with the original report. This means that for having an Action, Inaction, and Control norms, when considering the past behaviour {Action was taken, or Inaction}, we can arrive at the mean and the standard deviation between the two categories.

This paper will make use of both experiment 2 and experiment 3 and their provided data. These population groups are specifically MTurk hires who are separate from each other, i.e. experiment 2 participants are not allowed to participate in experiment 3. 

Experiment 2 consisted of 300 American hires, with a mean age of 36.58. This experiment incorporated a use of a control in order to make comparisons without the influence of societal norms.

Experiment 3 consisted of 303 American hires, with a mean age of 36.66. This experiment did not make use of social norms, but instead focused on expectations and past-behavioural tendencies. 
# Results


(@tbl-regretvalues)

# Discussion

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

```{r}
#| echo: false
#| warning: false
#| label: tbl-regretvalues
#| tbl-cap: "Perceived regret in relation to past behaviour"

exp2_table <- read_csv("../data/analysis_data/exp2_tibble.csv")

exp2_table |>
  kable(table.attr = "style='width:50%;'")
```

\newpage

# References
